---
title: "Divvy Ride Share"
author: "Deborah Luckey"
date: "7/6/2021"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(data.table)
 
```

```{r, include=FALSE}
#install.packages("readxl")
library(readxl)
library(knitr)
```
```{r, include=FALSE}
my_data <- read_excel("my_excel_file.xlsx")
```

```{r include=FALSE}
library(zoo)
```





## Agenda

- Purpose
- Analysis of data for casual riders and members
- Conclusion

## Purpose

To better understand patterns between casual riders and members and identify trends.

## Average Ride Length and Most Popular Day of Week

```{r convert to minutes, include=FALSE}
my_data$ride_length=as.numeric(my_data$ended_at - my_data$started_at, units="mins")
my_data$ride_length[1]

```

```{r get mode function, include=FALSE}

getmode <- function(v) {
   uniqv <- unique(v)
   uniqv[which.max(tabulate(match(v, uniqv)))]
}


```


```{r summary, echo=FALSE}
tripdata_summary <- my_data %>% 
  group_by(member_casual) %>% 
  summarise(avg_ride_length=round(mean(ride_length),2),
  popular_day=getmode(day_of_week))

  knitr::kable(tripdata_summary, "pipe", col.names = c("Rider Type", "Avg Ride Length", "Popular Day of Week"), align = c("l", "c", "c"))
```

Average ride length is in minutes.  
Day of week is numeric starting with Sunday as 1.

## Popular Day of Week by Rider Type

```{r include=FALSE}
avg_by_day_of_week <- my_data %>% 
  group_by(member_casual, day_of_week) %>% 
  summarise(avg_ride_length=mean(ride_length))

```


```{r day of week}
ggplot(data=avg_by_day_of_week, aes(x=day_of_week, y=avg_ride_length, fill=member_casual)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  labs(caption = "Average ride length is in minutes.  
Day of week is numeric starting with Sunday as 1.")

```





## Length of Ride by Season

```{r seasons, include=FALSE}
yq <- as.yearqtr(as.yearmon(my_data$started_at, "%m/%d/%Y") + 1/12)
my_data$Season <- factor(format(yq, "%q"), levels = 1:4, 
                    labels = c("winter", "spring", "summer", "fall"))

```

```{r season summary, include=FALSE}
season_summary <- my_data %>% 
  group_by(member_casual, Season) %>% 
  summarise(avg_ride_length=mean(ride_length))

```

```{r plot}
ggplot(season_summary, aes(x=Season, y=avg_ride_length, group=member_casual)) +
  geom_line(aes(linetype=member_casual, color=member_casual))+
  geom_point(aes(color=member_casual))+
  theme(legend.position="top")+
  labs(caption = "Day of week is numeric starting with Sunday as 1")

```


## Conclusion

- The casual riders average length time on the bike is double that of the member riders.
- Though both riders ride more on weekend, casual riders have a greater average length time.
- Both riders spend more time on bikes in summer, but casual riders average time is considerably more.
